Lesson 3
========================================================

***

### What to Do First?
Notes:

***

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}
setwd("/Users/michellepetersen/Downloads/eda-course-materials/lesson3")
list.files()
pf <- read.delim('pseudo_facebook.tsv')
names(pf)
dim(pf)
```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
install.packages('ggplot2')
library(ggplot2)
install.packages('ggthemes', dependencies = TRUE)
library(ggthemes)

qplot(x = dob_day, data = pf) +
  scale_x_continuous(breaks = 1:31)

ggplot(aes(x = dob_day), data = pf) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = 1:31)
```

***

#### What are some things that you notice about this histogram?
Response:

***

### Moira's Investigation
Notes:

Does their audience match thier perceived audience.
Estimated people's audience size.

***

### Estimating Your Audience Size
Notes:

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response: london sunset

#### How many of your friends do you think saw that post?
Response: 60%

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response: 50%

***

### Perceived Audience Size
Notes:

***
### Faceting
Notes:

```{r Faceting}
qplot(x = dob_day, data = pf) +
  scale_x_continuous(breaks = 1:31) +
  facet_wrap(~dob_month, ncol = 3)
#one variable

qplot(x = friend_count, data = pf, xlim = c(0, 1000))
#two variables
```

#### Let’s take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:


***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
qplot(x = friend_count, data = pf)
```

#### How is this plot similar to Moira's first plot?
Response:

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}
qplot(x = friend_count, data = pf, xlim = c(0, 1000))

qplot(x = friend_count, data = pf) +
  scale_x_continuous(limits = c(0,1000))

qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))

ggplot(aes(x = friend_count), data = pf) +
  geom_histogram(binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50))
```

### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_grid(gender ~ .)

ggplot(aes(x = friend_count), data = pf) +
  geom_histogram(binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
    facet_grid(gender ~ .)

ggplot(aes(x = friend_count), data = pf) +
  geom_histogram() +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), 
  binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)

ggplot(aes(x = friend_count), data = subset(pf, !is.na(gender))) +
  geom_histogram() +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
table(pf$gender)
by(pf$friend_count, pf$gender, summary)
```

#### Who on average has more friends: men or women?
Response:

#### What's the difference between the median friend count for women and men?
Response:

#### Why would the median be a better measure than the mean?
Response:

***

### Tenure
Notes:

```{r Tenure}

summary(pf$tenure)

qplot(x = tenure, data = pf, binwidth = 30,
      color = I('black'), fill = I('#099DD9'))

qplot(x = tenure/365, data = pf, binwidth = .25,
      color = I('black'), fill = I('#099DD9'))

ggplot(aes(x = tenure), data = pf) +
  geom_histogram(binwidth = 30, color = 'black', fill = '#099DD9')

ggplot(aes(x = tenure/365), data = pf) +
  geom_histogram(binwidth = .25, color = 'black', fill = '#F79420')

qplot(x = tenure/365, data = pf, binwidth = .25,
  color = I('black'), fill = I('#F79420')) +
  scale_x_continuous(breaks = seq(1, 7, 1), limits = c(0, 7))
```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}

```

***

### Labeling Plots
Notes:

```{r Labeling Plots}
ggplot(aes(x = tenure / 365), data = pf) +
  geom_histogram(color = 'black', fill = '#F79420') +
  scale_x_continuous(breaks = seq(1, 7, 1), limits = c(0, 7)) +
  xlab('Number of years using Facebook') +
  ylab('Number of users in sample')

qplot(x = tenure/365, data = pf, binwidth = .25,
  color = I('black'), fill = I('#F79420')) +
  scale_x_continuous(breaks = seq(1, 7, 1), limits = c(0, 7)) + 
  xlab('Number of years using Facebook') +
  ylab('Number of users in sample')

qplot(x = tenure/365, data = pf, binwidth = .25,
  xlab = 'Number of years using Facebook',
  ylab = 'Number of users in sample',
  color = I('black'), fill = I('#F79420')) +
  scale_x_continuous(breaks = seq(1, 7, 1), limits = c(0, 7)) 
```

***

### User Ages
Notes:

```{r User Ages}
summary(pf$age)

qplot(x = age, data = pf, binwidth = 1,
  xlab = 'Age of user using Facebook',
  ylab = 'Number of users in sample',
  color = I('black'), fill = I('#5760AB')) +
  scale_x_continuous(breaks =  seq(0, 113, 5)) 

ggplot(aes(x = age), data = pf) +
  geom_histogram(binwidth = 1, color = I('black'), fill = '#5760AB') +
  scale_x_continuous(breaks = seq(0, 113, 5)) +
  xlab('Age of users using Facebook') +
  ylab('Number of users in sample')

```

#### What do you notice?
Response:

***

### The Spread of Memes
Notes:
They tend to reoccur over time
***

### Lada's Money Bag Meme
Notes:
Money bag meme says it only occurs once in 800 years.  Log scale shows meme persistent in low numbers.  line geome.  recaling y axis to log

There are two main reasons to use logarithmic scales in charts and graphs. The first is to respond to skewness towards large values; i.e., cases in which one or a few points are much larger than the bulk of the data. The second is to show percent change or multiplicative factors.

A linear price scale is plotted on the side of the chart there is an equal distance between the prices, and each unit change on the chart is represented by the same vertical distance on the scale, regardless of what price level the asset is at when the change occurs.

A logarithmic price scale is plotted so that the prices in the scale are not positioned equidistantly; instead, the scale is plotted in such a way that two equal percent changes are plotted as the same vertical distance on the scale.

Read more: What is the difference between a logarithmic price scale and a linear one? https://www.investopedia.com/ask/answers/05/logvslinear.asp#ixzz5Vd0X8ebt 
Follow us: Investopedia on Facebook

https://www.investopedia.com/ask/answers/05/logvslinear.asp
***

### Transforming Data
Notes:

***

```{r Add a Transform}

install.packages('gridExtra')
library(gridExtra)

summary(pf$friend_count)

summary(log10(pf$friend_count + 1))

summary(sqrt(pf$friend_count))

qplot(x = friend_count, data = pf)
```

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}
# define individual plots
p1 = qplot(x = friend_count, data = pf, binwidth = 10)
p2 = qplot(x = log10(pf$friend_count + 1), data = pf)
p3 = qplot(x = sqrt(pf$friend_count), data = pf, binwidth = 5)
# arrange plots in grid
grid.arrange(p1, p2, p3, ncol=1)

p1gg = ggplot(aes(x = friend_count), data = pf) + geom_histogram()
p2gg = p1gg + scale_x_log10()
p3gg = p1gg + scale_x_sqrt()
grid.arrange(p1, p2, p3, ncol=1)
```

***


### Frequency Polygons
Allows us to compare distributions

```{r Frequency Polygons}

qplot(x = friend_count, y = ..count../sum(..count..),
    data = subset(pf, !is.na(gender)), 
  geom = "freqpoly", color = gender) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  xlab('Friend Count') +
  ylab('Proportion of users with that friend count')

ggplot(aes(x = friend_count, y = ..count../sum(..count..)),
       data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender), binwidth=10) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  xlab('Friend Count') +
  ylab('Proportion of users with that friend count')
```

***

### Likes on the Web
Notes:

```{r Likes on the Web}
qplot(x =www_likes, 
  data = subset(pf, !is.na(gender)), 
  geom = "freqpoly", color = gender) +
  scale_x_log10() + 
  xlab('Likes') +
  ylab('Count of users with that like count')

ggplot(aes(x = www_likes), 
  data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender)) +
  scale_x_log10() +
  xlab('Likes') +
  ylab('Count of users with that like count')

summary(by(pf$www_likes, pf$gender, sum))

```


***

### Box Plots
Notes:

```{r Box Plots}
by(pf$friend_count, pf$gender, summary)

qplot(x = gender, y = friend_count, 
    data = subset(pf, !is.na(gender)),
    geom = "boxplot") +
    coord_cartesian(ylim = c(0, 1000))

ggplot(aes(x = gender, y = friend_count), 
  data = subset(pf, !is.na(gender))) +
  geom_boxplot(aes(color = friend_count)) +
  coord_cartesian(ylim = c(0, 500))
```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}
qplot(x = gender, y = friend_count, 
    data = subset(pf, !is.na(gender)),
    geom = "boxplot") +
    coord_cartesian(ylim = c(0, 1000))

ggplot(aes(x = gender, y = friend_count), 
  data = subset(pf, !is.na(gender))) +
  geom_boxplot(aes(color = friend_count)) +
  coord_cartesian(ylim = c(0, 1000))
```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}
by(pf$friend_count, pf$gender, summary)

```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}
by(pf$friendships_initiated, pf$gender, summary)

qplot(x = gender, y = friendships_initiated, 
    data = subset(pf, !is.na(gender)),
    geom = "boxplot") +
    coord_cartesian(ylim = c(0, 300))

```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}
names(pf)

summary(pf$userid)

summary(pf$mobile_likes)

summary(pf$mobile_likes > 0)

pf$mobile_check_in <- NA

pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)

pf$mobile_check_in <- factor(pf$mobile_check_in)

summary(pf$mobile_check_in)

sum(pf$mobile_check_in == 1) / length(pf$mobile_check_in)

```

Response:

***

### Analyzing One Variable
Reflection:

Close look at individual variables
types
missing values
outliers
distribution

changes limits, bin widths, transformed.





***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!