Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

Scatter plots x axis actually saw the post.  how many they guessed.

If they guessed it would fall on the line

Stripes are guesses are regular numbers.

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
getwd()
setwd("/Users/michellepetersen/Downloads/eda-course-materials/lesson5")
pf = read.delim('pseudo_facebook.tsv')
library(ggplot2)
library(dplyr)

ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)

ggplot(aes(x = age, y = friend_count),
      data = subset(pf, !is.na(gender))) +
    geom_line(aes(color = gender), stat = 'summary', fun.y = median)

pf.fc_by_age_gender <- pf %>%
  filter(!is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(
      mean_friend_count = mean(friend_count),
      median_friend_count = median(friend_count),
      n = n()) %>%
  ungroup() %>%
  arrange(age)

head(pf.fc_by_age_gender, 6)
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
ggplot(aes(x = age, y = median_friend_count, color = gender),
      data = pf.fc_by_age_gender) +
    geom_line()
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
install.packages('reshape2')
library(reshape2)

spread(subset(pf.fc_by_age_gender, 
       select = c('gender', 'age', 'median_friend_count')), 
       gender, median_friend_count)

head(pf.fc_by_age_gender, 6)

pf.fc_by_age_gender.wide1 <- dcast(pf.fc_by_age_gender,
                              age ~ gender, 
                              value.var = 'median_friend_count')

head(pf.fc_by_age_gender.wide1, 10)

library(tidyr)
pf.fc_by_age_gender.wide2 <-
  subset(pf.fc_by_age_gender[c('age', 'gender', 'median_friend_count')],
         !is.na(gender)) %>%
  spread(gender, median_friend_count) %>%
  mutate(ratio = female / male)

head(pf.fc_by_age_gender.wide2)

```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
ggplot(aes(x = age, y = female / male),
      data = pf.fc_by_age_gender.wide2) +
    geom_line() + 
    geom_hline(yintercept = 1, alpha = 0.5, linetype = 2, color = "orange")
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
pf$year_joined <- floor(2014 - pf$tenure/365)

head(pf, 10)
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
summary(pf$year_joined)
table(pf$year_joined)

pf$year_joined.bucket <- cut(pf$year_joined, 
                             c(2004, 2009, 2011, 2012, 2014))
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
table(pf$year_joined.bucket, useNA = 'ifany')

ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket))) +
  geom_line(aes(color = year_joined.bucket), 
            stat = 'summary', fun.y = median)

```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket))) +
  geom_line(aes(color = year_joined.bucket), 
            stat = 'summary', fun.y = mean) + 
  geom_line(linetype = 2, color = "black", stat = 'summary', fun.y = mean)
```

***

### Friending Rate
Notes:

```{r Friending Rate}
with(subset(pf, tenure >= 1), summary(friend_count / tenure))
```

***

### Friendships Initiated
Notes:

What is the median friend rate?
.22

What is the maximum friend rate?
417

```{r Friendships Initiated}
ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket), 
            stat = 'summary', fun.y = mean) 
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure > 1)) +
  geom_smooth(aes(color = year_joined.bucket))

```

***

### Sean's NFL Fan Sentiment Study
Notes:
positive to negative words ratio every 5 minutes - plotted standard error
moving average for every 5 minutes
aggregated up to one day started to see some signals
predicts sentiment as a function of time

looking for a natural spline + 7 day moving average

looking for a seven day moving average told the characteristics of the season


***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yo <- read.csv("yogurt.csv")
str(yo)

#Convert id from int to factor
yo$id <- factor(yo$id)
str(yo)

ggplot(aes(x = price), data = yo) +
  geom_histogram(binwidth = 5, color = 'black', fill = '#F79420') 

#setting a binwidth to 10 leads to a biased model and leads to 
# missing the blank spots


```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
summary(yo)
length(unique(yo$price))
table(yo$price)

yo <- transform(yo, all.purchases = strawberry + blueberry + pina.colada + plain + mixed.berry)
   
summary(yo$all.purchases)  

# or 
#yo$all.purchases <- yo$strawberry + ....
```

***

### Prices over Time
Notes:

```{r Prices over Time}
#mine
ggplot(aes(x = time, y = price), data = yo) + 
    geom_point() + 
    geom_smooth(method = 'lm', color = 'red')

#answer
ggplot(aes(x = time, y = price), data = yo) + 
    geom_jitter(alpha = 1/4, shape = 21, fill = I("#f79420")) 
  
```

***

### Sampling Observations
Notes:
multiple observation of same unit.  work with a small sample households in more detail.
analysis of a sub sample

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(4230)
sample.ids <- sample(levels(yo$id), 16)

ggplot(aes(x = time, y = price),
      data = subset(yo, id %in% sample.ids)) +
      facet_wrap( ~ id) +
      geom_line() +
      geom_point(aes(size = all.purchases), pch = 1)

set.seed(42)
sample.ids <- sample(levels(yo$id), 16)

ggplot(aes(x = time, y = price),
      data = subset(yo, id %in% sample.ids)) +
      facet_wrap( ~ id) +
      geom_line() +
      geom_point(aes(size = all.purchases), pch = 1)

#Faceting over the households over time
# facet over the primary unit, case, or individual in the dataset over time.


```

***

### The Limits of Cross Sectional Data
Notes:

The facebook data is one snapshot of a characteristic of the individual.  Not the individual over time.

***

### Many Variables
Notes:

Much of the analysis so far was some pre-chosen variable, relationship, or question of interest and then used EDA to let the data speak.  Then incorporated a third or fourth variable to plot.  Summarize small set of variables into smaller dimensions.

Speed up EDA.  make many plots at once.

***

### Scatterplot Matrix
Notes:
```{r}
install.packages('GGally')
library(GGally)
theme_set(theme_minimal(20))

set.seed(1836)
pf_subset <- pf[, c(2:15)]
#or
pf_subset = pf[, c('age', 'dob_year', 'dob_month', 'gender', 'tenure')]
#or
pf_subset <- subset(pf, select = -c(userid, year_joined, year_joined_bucket))

names(pf_subset)
ggpairs(pf_subset[sample.int(nrow(pf_subset), 1000), ])

```

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)

```

```{r}
library(reshape2)
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

