Prosper Loan Exploration by Michelle Petersen
========================================================

#Overview 
This report explores the Prosper loans dataset. According to Wikipedia, 
Prosper Marketplace is a company in the peer-to-peer lending industry and 
operates 
[Prosper.com.](https:www.Prosper.com) Borrowers request personal loans on 
Prosper and investors can fund them. Prosper lists, collects and distributes 
borrower payments and interest back to the loan investors. 

The dataset contains a little over 113,000 loans with 81 variables on each 
loan. Click this link for list of all the variable definitions: 
[Prosper Loan Data - Variable Definitions.](https://docs.google.com/spreadsheets/d/1gDyi_L4UvIrLTEC6Wri5nbaMmkGmLQBk-Yx3z0XDEtI/edit#gid=0)

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

knitr::opts_chunk$set(fig.width=9,fig.height=5,fig.path='Figs/',
                      fig.align='center',tidy=TRUE,
                      echo=FALSE,warning=FALSE,message=FALSE)

suppressMessages(library(ggplot2))
suppressMessages(library(GGally))
suppressMessages(library(dplyr))
suppressMessages(library(tidyr))
suppressMessages(library(lubridate))
suppressMessages(library(gridExtra))
suppressMessages(library(memisc))
suppressMessages(library(pastecs))
suppressMessages(library(ggthemes))
suppressMessages(library(RColorBrewer))
```

```{r echo=FALSE, Load_the_Data}
setwd("/Users/michellepetersen/DataAnalyst/ExploratoryDataAnalysisInR")
prosper_loan_dataset <- read.csv('prosperLoanData.csv')
```

```{r echo=FALSE, Preliminary_Exploration_Dataset_Transformation}
# Convert the date from a factor to a Date
prosper_loan_dataset$LoanOriginationDate <- 
  strptime(as.character(prosper_loan_dataset$LoanOriginationDate), 
           format = "%Y-%m-%d %H:%M:%S")
prosper_loan_dataset$LoanOriginationDate <- 
  as.Date(as.character(prosper_loan_dataset$LoanOriginationDate), 
          format = "%Y-%m-%d")

# Create a column for the origination year
prosper_loan_dataset$OriginationYear <- 
  year(as.Date(as.character(prosper_loan_dataset$LoanOriginationDate), 
               format = "%Y-%m-%d"))

# Create a column for the origination month
prosper_loan_dataset$OriginationMonth <- 
  month(as.Date(as.character(prosper_loan_dataset$LoanOriginationDate), 
                format = "%Y-%m-%d"))

# Create a column for the origination day
prosper_loan_dataset$OriginationDay <- 
  day(as.Date(as.character(prosper_loan_dataset$LoanOriginationDate), 
              format = "%Y-%m-%d"))

# Add order to the quarter column so it sorts correctly by year and 
# quarter eg Q1 2009
prosper_loan_dataset$LoanOriginationQuarter = 
  factor(prosper_loan_dataset$LoanOriginationQuarter, 
    levels = unique(prosper_loan_dataset$LoanOriginationQuarter[
              order(prosper_loan_dataset$OriginationYear, 
                    prosper_loan_dataset$OriginationMonth)]),
              ordered=TRUE)

# Create a factor column for loans pre-2010 and post-2010 to facilitate 
# plotting
prosper_loan_dataset$OriginationBefore2010 <- 
  ifelse(prosper_loan_dataset$OriginationYear < 2010, 
         "pre-2010", "post-2010")
prosper_loan_dataset$OriginationBefore2010 <- 
  as.factor(prosper_loan_dataset$OriginationBefore2010)

# Convert ListingCategory to factor with strings rather
# than numbers for clarity
prosper_loan_dataset$ListingCategory <- 
  factor(prosper_loan_dataset$ListingCategory,
    levels =c (0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 
               11, 12, 13, 14, 15, 16, 17, 18, 19, 20),
    labels =c ("NotAvailable", "DebtConsolidation", "HomeImprovement",
              "Business", "PersonalLoan", "StudentUse", "Auto", "Other",
              "BabyAndAdoption", "Boat", "CosmeticProcedure", 
              "EngagmentRing",
              "GreenLoans", "HouseholdExpenses", 
              "LargePurchases", "MedicalDental",
              "Motorcycle", "RV", "Taxes", "Vacation", 
              "WeddingLoan"))

# Create another column as a factor for Credit Score values using
# common credit bureau terminology
prosper_loan_dataset$LowerCreditScoreGroup <- 
  cut(prosper_loan_dataset$CreditScoreRangeLower,
      include.lowest = FALSE,   
      breaks = c(0, 550, 649, 699, 749, 850),
      labels = c("Bad", "Poor", "Fair","Good", "Exceptional"))

# Create another column as a factor for Credit Score values using
# common credit bureau terminology
prosper_loan_dataset$UpperCreditScoreGroup <- 
  cut(prosper_loan_dataset$CreditScoreRangeUpper,
      include.lowest = FALSE,
      breaks = c(0, 550, 649, 699, 749, 850),
      labels = c("Bad", "Poor", "Fair", "Good", "Exceptional"))

# Order the Income Range factor column to sort correctly
prosper_loan_dataset$IncomeRange <- 
  ordered(prosper_loan_dataset$IncomeRange, 
          levels = c("Not displayed", "Not employed", "$0", 
                    "$1-24,999", "$25,000-49,999",
                    "$50,000-74,999", "$75,000-99,999", "$100,000+"))

# Order the Loan Status factor column to sort correctly
prosper_loan_dataset$LoanStatus <- 
  ordered(prosper_loan_dataset$LoanStatus, 
          levels = c("Cancelled", "Chargedoff", "Completed", 
                      "Current", "Defaulted", "FinalPaymentInProgress",
                      "Past Due (1-15 days)", "Past Due (16-30 days)", 
                      "Past Due (31-60 days)", "Past Due (61-90 days)",
                      "Past Due (91-120 days)", "Past Due (>120 days)"))

badStandingGroup = c("Chargedoff", "Defaulted",
                     "Past Due (31-60 days)", "Past Due (61-90 days)",
                     "Past Due (91-120 days)", "Past Due (>120 days)")

# Create a new factor variable to aggregate Loan Status values that may
# affect a borrowers credit score
prosper_loan_dataset$LoanStatusGroup <- 
  as.factor(ifelse(prosper_loan_dataset$LoanStatus %in% badStandingGroup, 
                   "InDefault", "GoodStanding"))

# Order the Prosper Rating and Credit Grade factor columns to sort correctly
prosper_loan_dataset$ProsperRating..Alpha <- 
  ordered(prosper_loan_dataset$ProsperRating..Alpha, 
          levels = c("N/A", "HR", "E", "D", "C", "B", "A", "AA"))
prosper_loan_dataset$CreditGrade <- 
  ordered(prosper_loan_dataset$CreditGrade, 
          levels = c("N/A", "NC", "HR", "E", "D", "C", "B", "A", "AA"))

# Create two data frames for pre and post 2010 loans for easier graphing
pre_2010_loans <- subset(prosper_loan_dataset, 
                         OriginationBefore2010 %in% c("pre-2010"))
post_2010_loans <- subset(prosper_loan_dataset, 
                          OriginationBefore2010 %in% c("post-2010"))
```

#Univariate Plots Section 

```{r echo=FALSE, messages=FALSE, Univariate_Plots_LoanOriginationYear}
# Add a bin width of 30 so plot is each month.  Gap in 2009 is clearly seen.
ggplot(aes(x = LoanOriginationDate), data = prosper_loan_dataset) +
  geom_histogram(binwidth = 30) +
  scale_x_date(date_breaks = "years", date_labels = "%Y") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Loan Origination Count by Year")
```

####Analysis of Loan Origination Count by Year
There is a surprising drop in loans in 2009.  In searching the news during 
that time I found this [Prosper News Story](http://archive.fortune.com/2009/04/28/technology/hempel_prosper.fortune/index.htm).  
In October of 2009, the SEC forced Prosper.com, to stop brokering new loans 
temporarily while it determined whether Prosper's loans should be classified 
as securities. After a six month quiet period Prosper was reopened to lenders 
and borrowers.  Prosper made other changes for its business including only 
allowing borrowers with a credit score above 640 to request a loan.  I added 
cohort for before 2010 and after 2010 to see other interesting similarities 
and differences. 

```{r echo=FALSE, message=FALSE, Univariate_Plots_LoanOriginationQuarter}
# Plot count of loans by quarter
ggplot(aes(fill = OriginationBefore2010, 
           x = LoanOriginationQuarter, ..count..), 
          data = prosper_loan_dataset) + 
  geom_bar() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Loan Origination Count by Quarter")
```

####Analysis of Loan Origination Count by Quarter
I've added color differences in the graphs for before and after the re-open  
in 2010.  There is a drop at the beginning of 2013 followed by fairly  
consistent growth in number of loans peaking at 14450 in Q4 of 2013. 

```{r echo=FALSE, message=FALSE, Univariate_Plots_LoanOriginationMonth}
# Use binwidth of 1 and breaks 1:12 to see all data by month
ggplot(aes(x = OriginationMonth, fill = OriginationBefore2010), 
       data = prosper_loan_dataset) + 
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = 1:12) +
  ggtitle("Loan Origination Count by Month")
```

####Analysis of Loan Origination Count by Month
January is the biggest month for new loans, followed by October and December. 

```{r echo=FALSE, message=FALSE, Univariate_Plots_LoanOriginationDay}
# Use binwidth of 1 and breaks 1:31 to see all data by day
ggplot(aes(x = OriginationDay, fill = OriginationBefore2010), 
       data = prosper_loan_dataset) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = 1:31) +
  ggtitle("Loan Origination Count by Day of the Month")
```

####Analysis of Loan Origination Count by Day of the Month
The number of loans listed rises over the month and peaks towards the end of  
the month. The highest number appear the last day of the month on the 30th.

```{r echo=FALSE, message=FALSE, Univariate_Plots_LoanOriginalAmount}
# Use binwidth of 1000 because loan amounts are in increments of $1000
ggplot(aes(fill = OriginationBefore2010, x = LoanOriginalAmount), 
       data = prosper_loan_dataset) +
    geom_histogram(binwidth = 1000)  +
    ggtitle("Loan Count by Loan Original Amount")
```

#### Summary statistics for Loan Original Amount pre-2010
```{r echo=FALSE, message=FALSE, Univariate_Summary_LoanOriginalAmountPre}
summary(pre_2010_loans$LoanOriginalAmount)
```

#### Summary statistics for Loan Original Amount post-2010
```{r echo=FALSE, message=FALSE, Univariate_Summary_LoanOriginalAmountPost}
summary(post_2010_loans$LoanOriginalAmount)
```

####Analysis of Loan Original Amount
There is a difference in loan amount before 2010 and after 2010.  The minimum  
loan amount pre-2010 and post-2010 remained the same at $1000. However, the  
maximum loan amount increased from $25000 to $35000.  

The mean increased from $6050 to $9190.  Post-2010 there are spikes in number  
of loans at $4000, $10000, and $15000. 

For both pre-2010 and post-2010 data the distribution is skewed to the right  
and a small number of loans greater than $25000.

```{r echo=FALSE, message=FALSE, Univariate_Plots_ListingCategory}
# Calculate percentage for y value
ggplot(aes(x = ListingCategory, y = 100 * (..count../sum(..count..)), 
           fill = OriginationBefore2010),
       data = subset(prosper_loan_dataset)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab("Percentage") +
  labs(title = "Percentage of Loans by Listing Category") 
```

####Analysis of Listing Category
Pre-2010 the listing category wasn't very informative. Almost 17000 of the 
loans had a listing category of "Not Available".  Post-2010 since there are  
only 20 loans where this information is not available. 

The debt consolidation category is the largest with a little over 50% of the  
loans listed.

```{r echo=FALSE, message=FALSE, Univariate_Plots_MonthlyLoanPayment}
# Calculate count for y value. No need for transformation since max loan 
# amount is set at $30000
ggplot(aes(fill = OriginationBefore2010, x = MonthlyLoanPayment, ..count..),
       data = prosper_loan_dataset) +
    geom_histogram() +
    ggtitle("Loan Count by Monthly Loan Payment") 
```

#### Summary statistics for MonthlyLoanPayment pre-2010
```{r echo=FALSE, message=FALSE, Univariate_Summary_MonthlyLoanPaymentPre}
summary(pre_2010_loans$MonthlyLoanPayment)
```

#### Summary statistics for MonthlyLoanPayment post-2010
```{r echo=FALSE, message=FALSE, Univariate_Summary_MonthlyLoanPaymentPost}
summary(post_2010_loans$MonthlyLoanPayment)
```

####Analysis of Monthly Loan Payment
The distribution for MonthlyLoanPayment is right skewed - Both pre-2010 and  
post-2010 mean values are greater than the median and there is a long tail to  
the right. The median payments are $100 higher (256.40) post-2010 than  
pre-2010 (149.03).

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_Term}
# Use percentage for y value
ggplot(aes(x = Term, y = 100 * (..count../sum(..count..))),
       data = prosper_loan_dataset) +
    geom_bar( ) +
    ylab("Percentage") + 
    scale_x_continuous(breaks=c(12, 36, 60)) +
    ggtitle("Percentage of Loans by Term") 
```

####Analysis of Term
Loans are for 1 year, 3 year or 5 years Terms. Almost 80% of loans are for a  
3 year term.

```{r echo=FALSE, message=FALSE, Univariate_Plots_LowerCreditScoreGroup}
# Use count for y value
ggplot(data = prosper_loan_dataset, 
       aes(fill = OriginationBefore2010, x = LowerCreditScoreGroup, ..count..)) +
    geom_bar() +
    ggtitle("Loan Count by Lower Credit Score")
```

#### Summary statistics for Lower Credit Score pre-2010
```{r echo=FALSE, message=FALSE, Univariate_Summary_CreditScoreRangeLowerPre}
summary(pre_2010_loans$CreditScoreRangeLower)
```

#### Summary statistics for Lower Credit Score post-2010
```{r echo=FALSE, message=FALSE, Univariate_Summary_CreditScoreRangeLowerPost}
summary(post_2010_loans$CreditScoreRangeLower)
```

####Analysis of Lower Credit Score
Here are the Credit Score Ranges as defined by [experian, transunion, and equifax:](https://www.creditsesame.com/blog/credit/credit-score-range-for-experian-transunion-equifax/)

* Bad 0 - 550
* Poor 551 - 649
* Fair 650 - 699
* Good 700 - 749
* Excellent 750 and above

Pre-2010 loans were allowed to be listed with Proper for Borrowers with a  
"Bad" credit rating. In addition there were 591 pre-2010 loans listed where  
the lower credit score was "Not Available". 

The median rose from 640 pre-2010 to 700 post-2010.  For pre-2010 lower credit  
score data is skewed slightly to the right since the mean (648.2) is greater  
than the median (640).  The post-2010 lower credit score data is normally  
distributed - The median and mean are very close at 700 and 699 respectively.

```{r echo=FALSE, message=FALSE, Univariate_Plots_UpperCreditScoreGroup}
# Use count for y value
ggplot(data = prosper_loan_dataset, 
       aes(fill = OriginationBefore2010, 
           x = UpperCreditScoreGroup, ..count..)) +
    geom_bar() +
    ggtitle("Loan Count by Upper Credit Score")
```

#### Summary statistics for Upper Credit Score pre-2010
```{r echo=FALSE, message=FALSE, Univariate_Summary_CreditScoreRangeUpperPre}
summary(pre_2010_loans$CreditScoreRangeUpper)
```

#### Summary statistics for Upper Credit Score post-2010
```{r echo=FALSE, message=FALSE, Univariate_Summary_CreditScoreRangeUpperPost}
summary(post_2010_loans$CreditScoreRangeUpper)
```

####Analysis of Upper Credit Score
Pre-2010 borrowers had a upper credit score median of 659 while post-2010  
borrowers had a median upper credit score of 719. The upper credit score  
values look normally distributed.

```{r echo=FALSE, message=FALSE, Univariate_Plots_CreditGrade}
# Use percentage for y value and elminate NA for Credit Grade
ggplot(aes(x = CreditGrade, y = 100 * (..count../sum(..count..))),
       data = subset(pre_2010_loans, !is.na(CreditGrade))) +
  geom_bar() +
  labs(title = "Percentage of Loans by Credit Grade (pre-2010)") +
  ylab("Percentage")
```

####Analysis of Credit Grade
The Credit Grade is the rating that was assigned to pre-2010 loans at the  
time the listing went live. I could not determine via the Prosper website  
search how the Credit Grade was determined. Credit Grade, from lowest-risk  
to highest-risk, are labeled AA, A, B, C, D, E, HR ("High Risk"), and NC  
("No Credit").

There were 141 loans listed for No Credit "NC" borrowers with 3508 loans  
listed for "HR" - High Risk borrowers. It is surprising how many loans were  
listed for borrowers in the Credit Grade groups lower than C.

```{r echo=FALSE, message=FALSE, Univariate_Plots_ProsperRatingAlpha}
# Use percentage for y value and elminate NA for Prosper Rating
ggplot(aes(x = ProsperRating..Alpha, y = 100 * (..count../sum(..count..))),
       data = subset(post_2010_loans, !is.na(ProsperRating..Alpha))) +
  geom_bar() +
  ylab("Percentage") +
  labs(title = 'Percentage of Loans by Prosper Rating (post-2010)') 
```

####Analysis of Prosper Rating
Prosper Ratings, from lowest-risk to highest-risk, are labeled AA, A, B, C, D,  
E, and HR ("High Risk").  Post-2010 Prosper provides a proprietary "Prosper  
Rating" based on the company's estimation of that borrower's "estimated loss  
rate." 

According to the company, the Prosper Rating is determined by two scores:

* the credit score, obtained from an official credit reporting agency
* the Prosper Score, figured in-house based on the Prosper population

Even though Credit Rating and Prosper Rating look similar, I decided not to  
combine them for my analysis since there was a decision by the company to no  
longer use Credit Grade and switch to Prosper Rating.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots_ProsperScore}
# Use percentage for y value and elminate NA for Prosper Score
ggplot(aes(x = ProsperScore, y = 100 * (..count../sum(..count..))),
       data = subset(post_2010_loans, !is.na(ProsperScore))) +
  geom_bar() +
  ylab("Percentage") + 
  labs(title = 'Percentage of Loans by Prosper Score (post-2010)') 
```

#### Summary statistics for Prosper Score post-2010
```{r echo=FALSE, message=FALSE, Univariate_Summary_ProsperScore}
summary(post_2010_loans$ProsperScore)
```

####Analysis of Prosper Score
The Prosper Score is a custom risk score for post-2010 loans built using  
historical Prosper data. The score ranges from 1-11, with 11 being the best,  
or lowest risk score. The Prosper score estimates the probability of a loan  
going “bad,” where “bad” is the probability of going 60+ days past due within  
the first twelve months from the date of loan origination. [Prosper Scorecard Link](https://www.prosper.com/plp/general-prosper_score/)

The Prosper Score looks normally distributed.

```{r echo=FALSE, message=FALSE, Univariate_Plots_IncomeRange}
# Use percentage for y for Income Range
ggplot(aes(x = IncomeRange, y = 100 * (..count../sum(..count..)), 
           fill = OriginationBefore2010),
       data = prosper_loan_dataset) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab("Percentage") +
  ggtitle('Percentage of Loans by Income Range') 
```

####Analysis of Income Range
The distribution of Income Ranges is left skewed with over 50% of the loans  
falling in the $25,000 - $75,000 income ranges.  Prior to 2010 almost 6% of  
loans didn't include borrower income range information.  This does not appear  
to be the case post-2010 where loans do contain borrower income range  
information.

```{r echo=FALSE, message=FALSE, Univariate_Plots_BorrowerRate}
# Use percentage for y for Borrower Rate and set ylim so comparison is clear
# across charts
p1 <- ggplot(aes(x = BorrowerRate, y = 100 * (..count../sum(..count..))),
       data = subset(pre_2010_loans, !is.na(BorrowerRate))) +
  geom_histogram() +
  ylab("Percentage") + 
  ylim(0, 15) +
  ggtitle('Percentage Loans by Borrower Rate (pre-2010)')

p2 <- ggplot(aes(x = BorrowerRate, y = 100 * (..count../sum(..count..))),
       data = subset(post_2010_loans, !is.na(BorrowerRate))) +
  geom_histogram() +
  ylab("Percentage") + 
  ylim(0, 15) +
  ggtitle('Percentage Loans by Borrower Rate (post-2010)')
  
grid.arrange(p1, p2, ncol=1)
```

#### Summary statistics for Borrower Rate pre-2010
```{r echo=FALSE, message=FALSE, Univariate_Summary_BorrowerAPRPre}
summary(pre_2010_loans$BorrowerRate)
```

#### Summary statistics for Borrower Rate post-2010
```{r echo=FALSE, message=FALSE, Univariate_Summary_BorrowerAPRPost}
summary(post_2010_loans$BorrowerRate)
```

####Analysis of Borrower Rate
The Borrower Rate is the Borrower's interest rate for this loan. The median  
Borrower Rate for pre-2010 loans was 17% which is lower than the median  
18.75% post-2010. The max Borrower Rate for pre-2010 loans was 50% vs 36%  
post-2010.  The interquartile range pre-2010 is 11.25% where post-2010 the  
interquartile range is 12.09%.  

It surprised me that Borrowers were paying a higher rate for a Prosper loan  
post-2010 compared to pre-2010 since interests rates for things like mortgages  
were lower. [Link to Historical Mortgage Rates.](https://www.valuepenguin.com/mortgages/historical-mortgage-rates)

```{r echo=FALSE, message=FALSE, Univariate_Plots_LoanStatusGroup}
# Use percentage for y for Loan Status and set y axis to increment by 10%
p1 <- ggplot(aes(x = LoanStatusGroup, y = 100 * (..count../sum(..count..))),
       data = subset(pre_2010_loans, !is.na(LoanStatusGroup))) +
  geom_bar() +
  ylab("Percentage") + 
  scale_y_continuous(breaks = seq(0, 100, 10), limits = c(0, 100)) +
  ggtitle('Percentage Loans by Loan Status (pre-2010)')

p2 <- ggplot(aes(x = LoanStatusGroup, y = 100 * (..count../sum(..count..))),
       data = subset(post_2010_loans, !is.na(LoanStatusGroup))) +
  geom_bar() +
  ylab("Percentage") +
  scale_y_continuous(breaks = seq(0, 100, 10), limits = c(0, 100)) +
  ggtitle('Percentage Loans by Loan Status (post-2010)')
  
grid.arrange(p1, p2, ncol=1)
```

####Analysis of Loan Status
I defined a factor variable for loans in "GoodStanding"" vs "InDefault" to  
group loans for further analysis.  I grouped loans as "InDefault" if the loan  
status might affect a borrowers credit rating.

Loans in the "InDefault" group:

* [Chargedoff](https://www.experian.com/blogs/ask-experian/charged-off-debt-must-still-be-repaid/)
* [Defaulted](https://www.valuepenguin.com/loans/what-does-it-mean-to-default-on-a-loan)
* Past Due (31-60 days)
* Past Due (61-90 days)
* Past Due (91-120 days)
* Past Due (>120 days)

Pre-2010 the percentage of loans "In Default" was about 35%.  Post-2010 about  
10% of loans listed were "InDefault"

# Univariate Analysis

### What is the structure of your dataset?
The prosper loans dataset contains 113937 rows and 81 variables.  There are 70  
numeric variables and 26 factor variables.  The dataset met the criteria for  
tidiness as defined  
[by Wikipedia here](https://en.wikipedia.org/wiki/Tidy_data).

* Each variable you measure should be in one column.
* Each different observation of that variable should be in a different row.
* There should be one table for each "kind" of variable.
* If you have multiple tables, they should include a column in the table that  
allows them to be linked.

### What is/are the main feature(s) of interest in your dataset?
The main areas I am focused on are how the change in Prosper Loans business  
affected their two customer groups - borrowers and lenders. 

* Borrower - What features are of interest and may impact how much I will pay  
to borrow money?
* Lender - What features are of interest and may impact if the loan will get  
paid back?

* LoanOriginationDate
* LoanStatus
* BorrowerRate
* CreditGrade
* ProsperRating
* ProsperScore
* IncomeRange
* CreditScoreRangeLower
* CreditScoreRangeUpper
* ListingCategory
* LoanOriginalAmount

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

* Term
* MonthlyLoanPayment
* IncomeVerifiable
* DebtToIncomeRatio
* BankcardUtilization
* EmploymentStatus

### Did you create any new variables from existing variables in the dataset?
Yes I created some new variables namely:

* OriginationYear
* OriginationMonth
* OriginationDay
* UpperCreditScoreGroup
* LowerCreditScoreGroup
* LoanStatusGroup

Each of these variables helped with creating graphs with a different or  
labeled groups.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

I performed some operations to adjust the data mainly for correctly ordering  
factor variables.

* LoanOriginationQuarter
* ListingCategory
* IncomeRange
* LoanStatus
* ProsperRating
* CreditGrade

# Bivariate Plots Section

```{r echo=FALSE, warning=FALSE, message=FALSE, Bivariate_Plots_CreditGrade}
# Use percentage for y axis for Credit Grade and set y axis so comparison
# is consistent across charts
ggplot(aes(x = CreditGrade, y = 100 * (..count../sum(..count..)), 
           fill = UpperCreditScoreGroup),
       data = subset(pre_2010_loans)) +
  geom_bar() +
  ylim(0, 25) +
  ggtitle("Percent Loans by Credit Grade (pre-2010)") + 
  ylab("Percentage") 
```

####Analysis of Credit Grade and Credit Score Group
This chart shows the relationship between Credit Grade and Upper Credit Score.  
There are a surprising number of loans for borrowers either "No Credit"" or  
are considered "High Risk".  It also seems odd that all the borrowers in the  
"High Risk" are not all in the same category namely "Bad".  That would seem  
confusing to a lender to have this group split when the E and D Credit Grades  
are much more uniform.

```{r echo=FALSE, warning=FALSE, message=FALSE, BV_ProsperRatingAlpha}
# Set percentage for y and set y axis so comparison is consistent across
# plots
ggplot(aes(x = ProsperRating..Alpha, y = 100 * (..count../sum(..count..)), 
           fill = UpperCreditScoreGroup),
       data = post_2010_loans) +
  geom_bar() +
  ylim(0, 25) +
  ggtitle("Percent Loans by Prosper Rating (post-2010)") + 
  ylab("Percentage") 
```

####Analysis of Prosper Rating and Credit Score Group
This chart shows the relationship between Prosper Rating and Upper  
Credit Score.  Post-2010 there are no borrowers with a "Bad" credit score.  It  
also seems odd that there are borrowers with a "Poor" through "Exception"  
Credit Rating in the High Risk category.  That would seem very confusing to a  
lender to have this mix of credit ratings across the prosper rating groups. It  
is very odd to have a "High Risk"" loan for a borrower with an "Exceptional"  
credit rating. 

Even with the mix of Credit Ratings within a Prosper Rating, there is a clear  
increase in better credit scores as the prosper rating values increase.

```{r echo=FALSE, warning=FALSE, message=FALSE, BV_BorrowerRatevsCreditGrade}
# Set y axis so comparison is consistent across plots 
ggplot(aes(x = CreditGrade, y = BorrowerRate, colour = LoanStatusGroup), 
       data = pre_2010_loans) +
    ggtitle("Borrower Rate vs Credit Grade") + 
    ylim(0, 0.4) +
    geom_boxplot() 
```


```{r echo=FALSE, warning=FALSE, message=FALSE, BV_BorrowerRatevsProsperRating}
# Set y axis so comparison is consistent across plots
ggplot(aes(x = ProsperRating..Alpha, y = BorrowerRate, colour = LoanStatusGroup), 
       data = post_2010_loans) +
    ggtitle("Borrower Rate vs Prosper Rating") + 
    ylim(0, 0.4) +
    geom_boxplot() 
```

####Analysis of Borrower Rate vs Credit Grade and Prosper Rating
These two graphs show the relationship between Borrower Rating and the  
pre-2010 Credit Grade and post-2010 Prosper Rating.  The Borrower Rate for  
Credit Grades "NC" through "E" were between 15% - 28%.  For Credit Grades D,  
C, B, A, and AA there is downward stair step.  

The Borrower Rate for Prosper Ratings post-2010 "HR" have a median value above  
30% which is higher. There is a clear downward stair step towards "AA" with  
the Borrower Rate median and interquartile range below 10%. 

```{r echo=FALSE, message=FALSE, Bivariate_Plots_LoanStatusvsCreditGrade}
# Set y axis to percentage of Credit Grade and Prosper Rating
ggplot(aes(x = CreditGrade, 
                y = 100 * (..count../sum(..count..)), 
                fill = LoanStatusGroup),
       data = subset(pre_2010_loans, !is.na(CreditGrade))) +
  geom_bar() +
  ylim(0, 25) + 
  ggtitle("Loan Status by Credit Grade (pre-2010)") +
  ylab("Percentage") 
```

```{r echo=FALSE, message=FALSE, Bivariate_Plots_LoanStatusvsProsperRating}
ggplot(aes(x = ProsperRating..Alpha, 
                y = 100 * (..count../sum(..count..)), 
                fill = LoanStatusGroup),
       data = subset(post_2010_loans, !is.na(ProsperRating..Alpha))) +
  geom_bar() +
  ylim(0, 25) + 
  ggtitle("Loan Status by Prosper Rating (post-2010)") +
  ylab("Percentage") 
```

``````{r echo=FALSE, warning=FALSE, message=FALSE, BV_LoanStatusvsProsperScore}
# Set y axis to percentage for Prosper Rating and make consistent ylim for easy
# comparison.  Set y axis to 25 to be consistent.
ggplot(aes(x = ProsperScore, 
           y = 100 * (..count../sum(..count..)), 
           fill = LoanStatusGroup),
       data = subset(post_2010_loans, !is.na(ProsperScore))) +
  geom_bar() +
  ggtitle("Loan Status by Prosper Score (post-2010)") +
  ylab("Percentage") +
  ylim(0, 25)
```

####Analysis of Loan Status vs Prosper Rating, Credit Grade and Prosper Score
These three graphs compare the loan status to each of the three variables  
measuring credit quality and risk.  The Credit Grade graph shows that the 35%  
of defaults in loans pre-2010 occured across all the Credit Grades.  I was  
surprised to see about 2% of loans in AA in default and made me question the  
validity of the rating. As I mentioned earlier in the report, I could not find  
details on how the Credit Grade was determined.  

The Prosper Rating graph shows a decrease in loans "InDefault"" as the Prosper  
Rating increases. This is more in line with what I would expect for a borrower  
loan rating.  

The Prosper Score is an internal scorecard and estimates the probability of a  
loan going “bad,” but only looks at the possibility of a loan going bad within  
the first year of the loan.  In the univariate analysis, we saw that almost  
80% of the loans are for three years. As we can see in the chart above, loans  
with a higher prosper score  six or above also had about a 1% default rate. If  
prosper updated thier model to look at longer timelines, this score could be  
more accurate and reduce the post-2010 10% default rate.

```{r echo=FALSE, messages = FALSE, warning=FALSE, MV_LoanAmountvsCreditGrade}
# Use geom_jitter and alpha of 1/4 because of the number of points and since
# Credit Grade is discrete
ggplot(aes(x = CreditGrade, 
           y = LoanOriginalAmount, 
           colour = as.factor(LoanStatusGroup)),
       data = pre_2010_loans) +
    geom_jitter(height = 0) +
    scale_color_brewer(type = "qual", palette = "Set1",
                       name="Loan Status") +
    ggtitle("Loan Original Amount vs Credit Grade")
```

```{r echo=FALSE, Multivariate_Plots_LoanOriginalAmountvsProsperRating}
# Use geom_jitter and alpha of 1/4 because of the number of points and since
# Prosper Rating is discrete
ggplot(aes(x = ProsperRating..Alpha, 
           y = LoanOriginalAmount, 
           colour = as.factor(LoanStatusGroup)),
       data = post_2010_loans) +
    geom_jitter(height = 0) +
    scale_color_brewer(type = "qual", palette = "Set1",
                       name="Loan Status") +
    ggtitle("Loan Original Amount vs Prosper Rating")
```

####Analysis of Loan Original Amount vs Prosper Rating and Credit Grade
I used a scatter plot with an alpha of 1/4 to plot Loan Original Amounts vs 
the Credit Grade and Prosper Rating. The graph of Credit Grade shows a larger 
number of defaults for loan amounts greater than $5000 for Credit Grades B and  
lower.  The higher loan amounts even across credit grades are more likely to  
be "InDefault" as is shown by the predominantly blue color.  

The Prosper Rating graph shows the much lower rates of default as we saw  
earlier.  It also shows lower Loan Amounts for borrowers in the High Risk  
group post-2010 compared to pre-2010.  Post-2010 shows much more restraint and  
consistency when loaning money to borrowers with lower ratings.

```{r echo=FALSE, messages = FALSE, warning=FALSE, MV_MonthlyPaymentvsGrade}
# Use sqrt for Monthly Loan Payment because of skew. Normalize by using the 
# loan amounts of $10000 since pre and post had a high number of loans listed.
# Set y axis to be consistent for comparison.
ggplot(aes(x = CreditGrade, 
           y = MonthlyLoanPayment, 
           colour = LoanStatusGroup), 
       data = subset(pre_2010_loans, LoanOriginationAmount = 10000)) +
    geom_boxplot() +
    ylab("MonthlyLoanPayment (sqrt)") + 
    scale_y_sqrt(limits = c(0, 1000)) +
    ggtitle("Monthly Loan Payment vs Credit Grade for Loans of $10000")
```

```{r echo=FALSE, messages = FALSE, warning=FALSE, MV_MonthlyvsProsperRating}
# Use sqrt for Monthly Loan Payment because of skew. Normalize by using the 
# loan amounts of $10000 since pre and post had a high number of loans listed.
# Set y axis to be consistent for comparison.
ggplot(aes(x = ProsperRating..Alpha, 
           y = MonthlyLoanPayment, 
           colour = LoanStatusGroup), 
       data = subset(post_2010_loans, (LoanOriginationAmount = 10000) &
                     !(is.na(ProsperRating..Alpha)) &
                      !(is.na(LoanStatusGroup)))) +
    geom_boxplot() +
    ylab("MonthlyLoanPayment (sqrt)") + 
    scale_y_sqrt(limits = c(0, 1000)) +
    ggtitle("Monthly Loan Payment vs Prosper Rating for Loans of $10000")
```

####Analysis of Monthly Loan Payment vs Prosper Rating and Credit Grade
For both plots, I only looked at Loans for $10000 to be able to compare Monthly  
Loan Payments pre-2010 and post-2010.  I took the sqrt of the Monthly Loan  
Payment to adjust for the skewness in the data.  The plot for Monthly Loan  
Payments post-2010 shows that interquartile range for payments across all  
Prosper Ratings are below $500. For pre-2010 loans there it was interesting  
to see that interquartile range that are in default were all above $500.  We  
can see from these charts that a monthly loan payment below $500 has an impact  
on keeping the loan in good standing.  

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \ 
investigation. How did the feature(s) of interest vary with other features in \
the dataset?
I summarized my findings after each plot or group of plot above.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?
The Monthly Loan Payment analysis was interesting. Having a payment amount 
that the borrower can make each month has a big influence on keeping the loan 
in good standing.

### What was the strongest relationship you found?
The strongest relationship I found was BorrowerRate and ProsperRating. It 
was interesting because if a borrower wanted to reduce the rate they paid it 
wasn't as transparent to figure out what actions they could take. The Prosper 
Rating is composed of the Credit Score and the Prosper Score.  If they want to 
increase thier credit score they can call the credit service and see details. 
For the Prosper Rating or Prosper Score it isn't as clear how those are 
calculated and they affect the rate the borrower pays.

# Multivariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, MV_MonthlyLoanPaymentvsTerm}
# Use sqrt for Monthly Loan Payment because of skew. Subset to loans in bad
# standing post-2010 since we would like to look forward for recommendations. 
# Subset to Prosper Score < 6. Set y axis to be consistent for comparison.
# Set the y axis to show discrete values for Term.
ggplot(aes(x = Term, y = MonthlyLoanPayment, 
           colour = as.factor(ProsperScore)), 
      data = subset(post_2010_loans, 
                     (post_2010_loans$LoanStatus %in% badStandingGroup) & (ProsperScore < 6))) +
    geom_jitter(height = 0) +
    geom_smooth(method = "glm", se = FALSE) +
    ylab("MonthlyLoanPayment (sqrt)") + 
    scale_y_sqrt(limits = c(0, 1000)) +
    scale_x_continuous(breaks=c(12, 36, 60)) +
    scale_color_brewer(type = "div", palette = "RdYlBu", 
                     name="Prosper\nScore") +
    ggtitle("Monthly Loan Payment vs Term for Loans In Default (post-2010)") 
```
```{r echo=FALSE, message=FALSE, warning=FALSE, MV_MonthlyLoanPaymentvsTermGood}
# Use sqrt for Monthly Loan Payment because of skew. Subset to loans in good
# standing post-2010 to look forward for recommendations. 
# Subset to Prosper Score < 6.  Ensure Prosper Score is encoded as a discrete
# variable.  Combine regression lines by using aes(group=1)
# Set y axis to be consistent for comparison.
ggplot(aes(x = Term, y = MonthlyLoanPayment, 
           colour = as.factor(ProsperScore)),
      data = subset(post_2010_loans, 
                    !(post_2010_loans$LoanStatus %in% badStandingGroup) 
                    & (ProsperScore < 6))) +
    geom_jitter(height = 0) +
    ylab("MonthlyLoanPayment (sqrt)") + 
    geom_smooth(method = "glm", se = FALSE) +
    scale_y_sqrt(limits = c(0, 1000)) +
    scale_x_continuous(breaks=c(12, 36, 60)) +
    scale_color_brewer(type = "div", palette = "RdYlBu", 
                       name="Prosper\nScore") +
    ggtitle("Monthly Loan Payment vs Term for Loans In Good Standing (post-2010)") 
```

```{r echo=FALSE, message=FALSE, warning=FALSE, MV_MonthlPaymentvsAmountBad}
# Use sqrt for Monthly Loan Payment because of skew. Subset to loans in bad
# standing. Subset to Prosper Score < 6. Ensure Prosper Score is encoded as a discrete
# variable.  Combine regression lines by using aes(group=1)
# Set y axis to be consistent for comparison.
ggplot(aes(x = LoanOriginalAmount, 
           y = MonthlyLoanPayment, 
           colour = as.factor(ProsperScore)), 
       data = subset(post_2010_loans, 
                     (post_2010_loans$LoanStatus %in% badStandingGroup) 
                     & (ProsperScore < 6))) +
    geom_point() +
    geom_smooth(method = "glm", se = FALSE) +
    scale_y_sqrt(limits = c(0, 1000)) +
    ylab("Monthly Loan Payment (sqrt)") + 
    ggtitle("Loan Amount vs Monthly Payment for Loans In Default (post-2010)") +
    scale_color_brewer(type = "div", palette = "RdYlBu", 
                       name="Prosper\nScore")
```

```{r echo=FALSE, message=FALSE, warning=FALSE, MV_AmountvsMontlyPaymentGood}
# Use sqrt for Monthly Loan Payment because of skew. Subset to loans in bad
# standing. Subset to Prosper Score < 6. Ensure Prosper Score is encoded as a discrete
# variable.  Combine regression lines by using aes(group=1)
# Set y axis to be consistent for comparison.
ggplot(aes(x = LoanOriginalAmount, y = MonthlyLoanPayment, 
           colour = as.factor(ProsperScore)), 
       data = subset(post_2010_loans, 
                     !(post_2010_loans$LoanStatus %in% badStandingGroup) 
                     & (ProsperScore < 6))) +
    geom_point() +
    geom_smooth(method = "glm", se = FALSE) +
    scale_y_sqrt(limits = c(0, 1000)) + 
    ylab("Monthly Loan Payment (sqrt)") + 
    scale_color_brewer(type = "div", palette = "RdYlBu", 
                       name="Prosper\nScore") + 
    ggtitle("Loan Amount vs Monthly Payment for Loans In Good Standing (post-2010)")
```

```{r echo=FALSE, message=FALSE, warning=FALSE, MV_BorrowerRateLinearModel}
set.seed(20022012)
subset_vars <- NULL

ColumnGroup <- c("BorrowerRate",
                "CreditGrade", 
                "ProsperRating..numeric.",
                "ProsperScore",
                "IncomeRange", 
                "CreditScoreRangeLower",
                "CreditScoreRangeUpper",
                "ListingCategory",
                "LoanOriginalAmount",
                "Term",
                "MonthlyLoanPayment",
                "IncomeVerifiable",
                "DebtToIncomeRatio",
                "BankcardUtilization",
                "EmploymentStatus",
                "LoanStatus",
                "OriginationBefore2010",
                "LoanStatusGroup")

subset_vars <- select_at(post_2010_loans, ColumnGroup)
subset_vars <- subset(subset_vars, !is.na(ProsperScore))

m1 <- lm(BorrowerRate ~ Term, data = subset_vars)
m2 <- update(m1, ~ . + CreditScoreRangeUpper)
m3 <- update(m2, ~ . + DebtToIncomeRatio)
m4 <- update(m3, ~ . + ProsperScore)
m5 <- update(m4, ~ . + CreditScoreRangeUpper)
m6 <- update(m5, ~ . + ProsperRating..numeric.)

mtable123456 <- mtable("Model 1 " = m1, "Model 2 " = m2, "Model 3 " = m3, 
                    "Model 4 " = m4, "Model 5 " = m5, "Model 6 " = m6,
                    summary.stats=c("sigma", "R-squared", "F", "p", "N"))
mtable123456
```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

I focused my analysis in this section on looking at how to increase the 
ability for borrowers to keep thier loan in good standing.  

The first two plots looked at MontlyLoanPayment vs Term by LoanStatus for 
borrowers with a ProsperScore below 6.  I thought about looking at this as 
if I were a data analyst for Prosper - There might be an opportunity for 
Prosper to increase the Term or provide more flexibility on the Term (4 year 
loans?) to reduce the monthly payment amount to < $250 and increase the 
ability for the borrower to make the lower payments.

The second two plots looked at MonthlyLoanPayment vs Loan Original Amount by 
Loan Status for borrowers with a ProsperCore below 6.  Again I thought about 
looking at this as if I were a data analyst for Prosper - There might be an 
opportunity to refine the Loan Consolidation model to keep the payment 
amounts lower.  The red line shows loans with the lowest prosper score and 
therefore most likely to go bad but shows a trend upwards for the payment.  If
the payments were lower perhaps the loans would be less risky.

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the 
strengths and limitations of your model.

I created a model for BorrowerRate to better understand how it was calculated. 
I looked at Term, CreditScoreRangeUpper, DebtToIncomeRatio, ProsperScore, and 
ProsperRating..numeric. The correlation between BorrowerRate and 
ProsperRating..numeric is strong at .92 R-squared. It was suprising the 
correlation for Credit Score was so low at .29.  It would be a challenge as a 
borrower to figure out how to get a better rate on the Prosper platform 
without increased transparency into how those metrics are calculated.

The models I put together are simplistic at this point compared to the number 
of variables associated with peer to peer lending but still insightful.  After 
I get more experience with model development beyond linear regression I would 
like to revisit this analysis. 

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_One}
p1 <- ggplot(aes(x = ProsperRating..Alpha, 
                 y = ..count../sum(..count..)),
       data = subset(post_2010_loans, 
                     !is.na(ProsperRating..Alpha) 
                     & !is.na(UpperCreditScoreGroup))) +
  geom_bar(aes(fill = LowerCreditScoreGroup)) +
  ggtitle("Loans vs Prosper Rating for Loans After 2010") + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) + 
  ylab("Loans (Percentage)") +
  xlab("Prosper Rating") +
  labs(fill = "Lower Credit Score")

p1 + theme_grey() 
```

### Description One
I chose this plot because it showed a difference between the Credit Score and 
the Prosper Rating.  It was suprising that a "Poor" credit score would end 
appear in the higher Prosper Ratings.

### Plot Two
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Two}
p1 <- ggplot(aes(x = CreditGrade, y = MonthlyLoanPayment), 
       data = subset(pre_2010_loans, 
                     !is.na(CreditGrade) & !is.na(UpperCreditScoreGroup))) +
    geom_boxplot(aes(colour = LoanStatusGroup)) +
    ylab("Monthly Loan Payment (sqrt)") + 
    xlab("Credit Grade") +
    scale_y_sqrt(limits = c(0, 1000), labels = scales::dollar_format(accuracy=1)) +
    labs(title = "Monthly Payment vs Credit Grade for Loans Before 2010") 

p1 + theme_economist() + 
  scale_colour_economist(name="Loan Status", 
                         labels=c("In Good Standing","In Default"))
```

### Description Two
For this plot I looked at loans before 2010.  I took the sqrt of the Monthly 
Loan Payment to adjust for the skewness in the data.  It was interesting to 
see the medians montly payment for loans in good standing and loans in default 
drift as the monthly payment rose.  From this chart your can see that a lower 
monthly payment has an impact on keeping the loan in good standing.

### Plot Three
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Three}
ggplot(aes(x = LoanOriginalAmount, y = MonthlyLoanPayment, 
           colour = as.factor(ProsperScore)), 
       data = subset(post_2010_loans, 
                     (post_2010_loans$LoanStatus %in% badStandingGroup 
                      & ProsperScore < 7))) +
    geom_point() +
    geom_smooth(method = "glm", se = FALSE) +
    scale_y_sqrt(limits = c(0, 2000), labels = scales::dollar_format(accuracy=1)) +
    ylab("Monthly Payment (sqrt)") + 
    xlab("Loan Amount") +
    labs(title = "Loan Amount vs Monthly Payment for Loan Defaults After 2010") +
    scale_color_brewer(type = "div", palette = "RdYlBu", 
                       name="Prosper\nScore")
```

### Description Three
For this plot I looked at loans originated after 2010.  The graph shows how 
currently the monthly payment is correlated to the Loan Amount rather than how 
if a borrower can consistently make monthly payments.  (The correlation 
between Montly Loan Payment and Loan Original Amount when calculated with 
cor.test was .91.) As loan amounts increase, monthly payments increase. 

Prosper already made an adjustment in 2010 to keep payments lower.  It would 
be intesting for them to model if adding more options for Term and Monthly 
payment amounts would futher reduce the post-2010 10% default rate.

========================================================

# Reflection

For this exploratory data analysis I decided to take on one of the larger 
data sets. With 81 variables to explore, it was easy to get off track and 
look into each of the variables. It took a bit to figure out which variables 
were important and I was afraid to miss an insight. I did alot of plots as a 
result and many of them are not included because they not that useful. I 
decided to keep my focus on questions I thought Prosper's customers would 
care about.

I was very surprised at the Peer to Peer business environment pre-2010 that 
led to action from the SEC.  I found it incredibly helpful to have domain 
knowledge for Peer to Peer Lending  and overview of the business - both 
provided much needed context to what I was seeing in the data.

The interesting models I would like to predict are more complex and outside 
the scope of EDA.  For example the Prosper Score is an internal scorecard and 
estimates the probability of a loan going “bad,” where “bad” is the 
probability of going 60+ days past due within the first twelve months from 
the date of loan origination. This isn't helpful for loans that are 3 year or 
5 year in term.  The lenders would probably like a better score to predict if 
it is likely that a loan will get paid off.  It would also be good to know if 
focusing on lower monthly payments would help keep the loans in good standing. 
The loans are not secured, Prosper has discontinued use of a secondary market, 
and the lender is bearing the risk.

What did go well is my EDA skills have improved.  I also have a much better 
understanding of when to apply a transform to a variable and what plots to use 
and in what circumstance. The example EDA projects were inspiring and I was 
able to use a few tips.  I am excited to take on machine learning.  I would 
like to develop a model that attempt to forecast what type of loans are likely 
to stay in good standing as a future project.

### Additional References
[https://www.prosper.com](https://www.prosper.com)

[Prosper Loans Wikipedia](https://en.wikipedia.org/wiki/Prosper_Marketplace)

[Practical Statistics for Data Scientists](https://www.amazon.com/Practical-Statistics-Data-Scientists-Essential/dp/1491952962/ref=sr_1_2?ie=UTF8&qid=1543353837&sr=8-2&keywords=practical+statistics+for+data+scientists)

[https://cran.r-project.org](https://cran.r-project.org)

[Prosper Score Development](https://www.prosper.com/plp/general-prosper_score/)

[StackOverflow Regression in R](https://stackoverflow.com/questions/18067519/using-r-to-do-a-regression-with-multiple-dependent-and-multiple-independent-vari)

[RDocumentation](https://www.rdocumentation.org/packages/scales/versions/0.2.0/topics/dollar_format)