Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

Scatter plots x axis actually saw the post.  how many they guessed.

If they guessed it would fall on the line

Stripes are guesses are regular numbers.

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
pf = read.delim('pseudo_facebook.tsv')

pf = read.csv('pseudo_facebook.tsv', sep = '\t')
```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
qplot(x= age, y= friend_count, data = pf)

qplot(age, friend_count, data = pf)

ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point() +
  xlim(13, 90)

summary(pf$age)
```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha = 1/20) +
  xlim(13, 90)

ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_jitter(alpha = 1/20) +
  xlim(13, 90)

```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha = 1/20, position = position_jitter(h = 0)) +
  xlim(13, 90) + 
  coord_trans( y = "sqrt")

```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha = 1/20, position = position_jitter(h = 0)) +
  xlim(13, 90) + 
  coord_trans( y = "sqrt")
```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}

summary(pf$friendships_initiated)

ggplot(aes(x = age, y = friendships_initiated), data = pf) + 
   geom_point(alpha = 1/15, position = position_jitter(h = 0)) + 
    xlim(13, 90) 
    coord_trans( y = "sqrt")
```

***

### Overplotting and Domain Knowledge
Notes:

population of the possible.  Use percentage

***

### Conditional Means
Notes:

filter()
group_by()
mutate()
arrange()

```{r Conditional Means}
# Without chaining
age_groups <- group_by(pf, age)

pf.fc_by_age <- summarise(age_groups,
    friend_count_mean = mean(friend_count),
    friend_count_median = median(friend_count),
    n = n())

pf.fc_by_age <- arrange(pf.fc_by_age, age)

head(pf.fc_by_age)

#Using dplyr
pf.fc_by_age <- pf %>%
  group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count),
    friend_count_median = median(friend_count),
    n = n()) %>%
  arrange(age)

head(pf.fc_by_age, 20)
    
```

Create your plot!

```{r Conditional Means Plot}
ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age) + 
   geom_line() 

```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x = age, y = friend_count), data = pf) + 
    xlim(13, 90) +
    geom_point(alpha = 0.05, 
               position = position_jitter(h = 0),
               color = 'orange') +
    coord_trans(y = 'sqrt') +
    geom_line(stat = 'summary', fun.y = mean,
              color = 'blue') +
    geom_line(stat = 'summary', fun.y = quantile, 
              fun.args = list(probs = .1),
              linetype = 2, color = 'blue') +
    geom_line(stat = 'summary', fun.y = quantile, 
              fun.args = list(probs = .9),
              linetype = 2, color = 'blue') +
    coord_cartesian(xlim = c(13, 90))

ggplot(aes(x = age, y = friend_count), data = pf) + 
    geom_point(alpha = 0.05, 
               position = position_jitter(h = 0),
               color = 'orange') +
    geom_line(stat = 'summary', fun.y = mean,
              color = 'blue') +
    geom_line(stat = 'summary', fun.y = quantile, 
              fun.args = list(probs = .1),
              linetype = 2, color = 'blue') +
    geom_line(stat = 'summary', fun.y = quantile, 
              fun.args = list(probs = .9),
              linetype = 2, color = 'blue') +
    coord_cartesian(xlim = c(13, 70), ylim = c(0, 1000))
  
```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:
Bernstein, M. S., Bakshy, E., Burke, M., & Karrer, B. (2013). Quantifying the invisible audience in social networks. In Proceedings of the SIGCHI Conference

http://hci.stanford.edu/publications/2013/invisibleaudience/invisibleaudience.pdfe on Human Factors in Computing Systems (CHI 2013), pp. 21-30.


***

### Correlation
Notes:

```{r Correlation}

cor.test(pf$age, pf$friend_count)

with(pf, cor.test(pf$age, pf$friend_count))

```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf, age <= 70), cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) + 
    xlim(0, 1000) +
    ylim(0, 2500) +
    geom_point(alpha = 0.05, 
               position = position_jitter(h = 0),
               color = 'orange') +
    geom_line(stat = 'summary', fun.y = mean,
              color = 'blue') +
    coord_trans(y = 'sqrt')

ggplot(aes(x = www_likes_received, y = likes_received), data = pf) + 
    xlim(0, quantile(pf$www_likes_received, 0.95)) +
    ylim(0, quantile(pf$likes_received, 0.95)) +
    geom_point() + 
    geom_smooth(method = 'lm', color = 'red')


```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
 cor.test(pf$www_likes_received, pf$likes_received)
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
 cor.test(pf$www_likes_received, pf$likes_received)
```

Response:

***

### Moira on Correlation
Notes:

Scatter plots to see visually how correlated two variables there are. Assumption of regression is that two are too highly correlated to one another.  Which to throw away and which to keep.

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)

data(Mitchell)
names(Mitchell)
```

Create your plot!

```{r Temp vs Month}
ggplot(aes(x = Month, y = Temp), data = Mitchell) + 
    geom_point() + 
    geom_smooth(method = 'lm', color = 'red')

ggplot(aes(x = Month, y = Temp), data = Mitchell) + 
    geom_point() + 
    geom_smooth(method = 'lm', color = 'red')

```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
cor.test(Mitchell$Month, Mitchell$Temp)
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
#Mine
ggplot(aes(x = Month, y = Temp), data = Mitchell) + 
    geom_point() + 
    scale_x_continuous(breaks = Mitchell$Month%%12, limits = c(1, 12))

#Theirs
ggplot(aes(x = Month, y = Temp), data = Mitchell) + 
    geom_point() + 
    scale_x_continuous(breaks = seq(0, 203, 12))

ggplot(aes(x = (Month%%12), y = Temp), data = Mitchell) +
  geom_point()

```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
library(energy)
dcor.test(Mitchell$Month, Mitchell$Temp, R = NULL)

pf$age_with_months <- pf$age + (12 - pf$dob_month / 12)

pf$age_with_months <- with(pf, age + (12 - dob_month / 12))

```

***

### Age with Months Means

```{r Age with Months Means}
age_groups_with_months <- group_by(pf, age_with_months)
pf.fc_by_age_months <- summarise(age_groups_with_months,
                          friend_count_mean = mean(friend_count),
                          friend_count_median = median(friend_count),
                          n = n())
pf.fc_by_age_months <- arrange(pf.fc_by_age_months, age_with_months)
head(pf.fc_by_age_months)

pf.fc_by_age_months <- pf %>%
  group_by(age_with_months) %>%
  summarise(friend_count_mean = mean(friend_count),
    friend_count_median = median(friend_count),
    n = n()) %>%
  arrange(age_with_months)

head(pf.fc_by_age_months, 20)

```

Programming Assignment
```{r Programming Assignment}
ggplot(aes(x = age_with_months, y = friend_count_mean), 
    data = subset(pf.fc_by_age_months, age_with_months < 71)) + 
    geom_line()
```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
p1 <- ggplot(aes(x = age, y = friend_count_mean), 
    data = subset(pf.fc_by_age, age < 71)) + 
    geom_line() + 
    geom_smooth()

p2 <- ggplot(aes(x = age_with_months, y = friend_count_mean), 
    data = subset(pf.fc_by_age_months, age_with_months < 71)) + 
    geom_line() + 
    geom_smooth()

p3 <- ggplot(aes(x = round(age / 5) * 5, y = friend_count), 
    data = subset(pf, age < 71)) + 
    geom_line(stat = 'summary', fun.y = mean)

library(gridExtra)
grid.arrange(p2, p1, p3, ncol = 1)
```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

